A simple and efficient algorithm for computing approximate Nash equilibria in two-person zero-sum sequential games with imcomplete information

### 2015-07-24 paper.tex
more

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polyhedron} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polyhedral strategy profiles $Q_k$
defined in \eqref{eq:polyhedron} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le j <
  n_1}{\text{max }}(Ay)_j - \underset{0 \le i < n_2}{\text{min
}}(A^Tx)_i.
\end{eqnarray}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplexes]{
    \includegraphics[width=.5\linewidth]{0.pdf}
  }
  %% \hspace{-2em}
  %% \subfigure[Toy Poker]{
  %%   \includegraphics[width=.5\linewidth]{SimplifiedPoker_dgap.pdf}
  %% }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.5\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-2em}
  \includegraphics[width=.5\linewidth]{1.pdf}
  %% \hspace{-1em}
  %% \includegraphics[width=.5\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.5\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: Evolution of ergodic
    primal-dual gap. \textbf{Bottom row}: Evolution of
    value of game...}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{Experiments on sequential games with imcomplete information: Kuhn 3-card
  poker.} This game is a simplified form of poker
developed by Harold W. Kuhn. It is a two-person zero-sum game which is
simply enough to serve as an example but contains all the complexity
traits of a imcomplete information sequential game. The
deck includes only three playing cards, for example a King, Queen, and
Jack. One card is dealt to each player, then the first player must bet
or pass, then the second player may bet or pass. If any player chooses
to bet the opposing player must bet as well ("call") in order to stay
in the round. After both players pass or bet the player with the
highest card wins the pot. The
sequence-form representation of the game has $n_1 = n_2 = 13$
sequences and $l_1 = l_2 = 7$ information
sets per player. More
 details (payoff matrix, etc.) are provided in the supplementary
 materials. The pair $(x^*, y^*) \in \mathbb{R}^{13} \times
\mathbb{R}^{13}$ of realization plans given by
\begin{eqnarray*}
  \begin{split}
    x^* &= [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0,
      0.275, 0.725]^T,\\
    y^* &= [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T
    \end{split}
\end{eqnarray*}
is a Nash $10^{-4}$-equlibrium computed in 1500 iterations of
Algorithm  \ref{Tab:algo}. The convergence curves are shown
in Fig \ref{Tab:dgap_curve}. One easy checks that this equilibrium is
feasible. Indeed, $E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67
      \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5},
      -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$
and
$E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
  \times 10^{-6}, -1.50 \times 10^{-6},
      2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T$.
Finally, one checks that ${x^*}^TAy^* = {-0.05555}$,
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of
 the game across iterations are shown in Figure \ref{Tab:dgap_curve}.
We have not benchmarked this against the algorithms proposed in
\cite{nesterov2005a} and Gilpin's et al. \cite{gilpinfirst} because
implementing them for such games would require us to compute the
complicated projections $\prox_{Q_k}$.  We recall that
avoiding these projections was one of the goals of the manuscript.

@@

\section{Supplementary materials}
\subsection{Sequence-form representation for Kuhn 3-card poker}
$E_1 \in \mathbb{R}^{7 \times 13}$ with $E_1(0,0) = E_1(1,9) =
E_1(1,12) = E_1(2,1) = E_1(2,4) = E_1(3,5) =
E_1(3,8) = E_1(4,2) = E_1(4,3) = E_1(5,6) = E_1(5,7) = E_1(6,10) =
E_1(6,11) = {1}$, $E_1(1,0) = E_1(2,0) = E_1(3,0) = E_1(4,1) =
E_1(5,5) = E_1(6,9) = {-1}$; \\

 $E_2 \in \mathbb{R}^{7 \times 13}$ with
$E_2(0,0) = E_2(1,7) = E_2(1,8) = E_2(2,9) = E_2(2,10) = E_2(3,5) =
E_2(3,6) = E_2(4,11) = E_2(4,12) = E_2(5,1) = E_2(5,2) = E_2(6,3) =
E_2(6,4) = {1}$, $E_2(1,0) = E_2(2,0) = E_2(3,0) = E_2(4,0) =
E_2(5,0) = E_2(6,0) = {-1}$; and \\

 $A \in \mathbb{R}^{13 \times
  13}$ with $A(3,8) = A(3,12) = A(4,6) = A(4,10) = A(7,12) = A(8,10) =
{-0.333333}$, $A(1,7) = A(1,11) = A(2,8) = A(2,12) = A(5,11) =
A(6,4) = A(6,12) = A(10,4) = A(10,8) = {-0.166667}$, $A(7,4) =
A(8,2) = A(11,4) = A(11,8) = A(12,2) = A(12,6) = {0.333333}$,
$A(4,5) = A(4,9) = A(5,3) = A(8,1) = A(8,9) = A(9,3) = A(9,7) =
A(12,1) = A(12,5) = {0.166667}$.\\

>>>>

\paragraph{A note about matrix games on simplexes.} It should be
noted
that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a matrix
  game with payoff matrix $A$, for which each player's strategy
profile is a simplex; this simplex can be written in the form
\eqref{eq:polyhedron} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game.  Thus the
results presented in this manuscript can be trivially applied such
games in particular. Here, the polyhedral strategy profiles $Q_k$
defined in \eqref{eq:polyhedron} reduce to simplexes $\Delta_{n_k}$,
and the primal-dual gap function $G(x,y)$ writes
\begin{eqnarray}G(x, y) =
\mathrm{max}\{\langle u, Ay\rangle - \langle x, Av\rangle | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\} = \underset{0 \le j <
  n_1}{\text{max }}(Ay)_j - \underset{0 \le i < n_2}{\text{min
}}(A^Tx)_i.
\label{eq:mg_pd}
\end{eqnarray}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplexes]{
    \includegraphics[width=.5\linewidth]{0.pdf}
  }
  %% \hspace{-2em}
  %% \subfigure[Toy Poker]{
  %%   \includegraphics[width=.5\linewidth]{SimplifiedPoker_dgap.pdf}
  %% }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.5\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-2em}
  \includegraphics[width=.5\linewidth]{1.pdf}
  %% \hspace{-1em}
  %% \includegraphics[width=.5\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.5\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: Evolution of ergodic
    primal-dual gap. \textbf{Bottom row}: Evolution of
    value of game. In \textit{(a)}, one can see the linear (i.e
    exponential) of the algorithm in \cite{gilpinfirst}, inbetween
    consecutive breakpoints on the $\epsilon$ grid. The duality gaps
    are computed according to formula \eqref{eq:mg_pd}. The
    first-order smoothing algorithm from \cite{nesterov2005a} jitters around as the
    iterations go on because even the smoothed problem becomes heavily
    ill-conditioned near solutions. On the other hand, our propsed
    algorithm seems to beat both of the aforementioned algorithms, and
  its proven $\mathcal{O}(1/\epsilon)$ convergence rate is clearly verified.}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{Experiments on sequential games with imcomplete information: Kuhn 3-card
  poker.} This game is a simplified form of poker
developed by Harold W. Kuhn. It is a two-person zero-sum game which is
simply enough to serve as an example but contains all the complexity
traits of a imcomplete information sequential game. The
deck includes only three playing cards, for example a King, Queen, and
Jack. One card is dealt to each player, then the first player must bet
or pass, then the second player may bet or pass. If any player chooses
to bet the opposing player must bet as well ("call") in order to stay
in the round. After both players pass or bet the player with the
highest card wins the pot. The
sequence-form representation of the game has $n_1 = n_2 = 13$
sequences and $l_1 = l_2 = 7$ information
sets per player. The sequence-form of the game is given by\\\\
$E_1 \in \mathbb{R}^{7 \times 13}$ with $E_1(0,0) = E_1(1,9) =
E_1(1,12) = E_1(2,1) = E_1(2,4) = E_1(3,5) =
E_1(3,8) = E_1(4,2) = E_1(4,3) = E_1(5,6) = E_1(5,7) = E_1(6,10) =
E_1(6,11) = {1}$, $E_1(1,0) = E_1(2,0) = E_1(3,0) = E_1(4,1) =
E_1(5,5) = E_1(6,9) = {-1}$; $E_2 \in \mathbb{R}^{7 \times 13}$ with
$E_2(0,0) = E_2(1,7) = E_2(1,8) = E_2(2,9) = E_2(2,10) = E_2(3,5) =
E_2(3,6) = E_2(4,11) = E_2(4,12) = E_2(5,1) = E_2(5,2) = E_2(6,3) =
E_2(6,4) = {1}$, $E_2(1,0) = E_2(2,0) = E_2(3,0) = E_2(4,0) =
E_2(5,0) = E_2(6,0) = {-1}$; and $A \in \mathbb{R}^{13 \times
  13}$ with $A(3,8) = A(3,12) = A(4,6) = A(4,10) = A(7,12) = A(8,10) =
{-0.333333}$, $A(1,7) = A(1,11) = A(2,8) = A(2,12) = A(5,11) =
A(6,4) = A(6,12) = A(10,4) = A(10,8) = {-0.166667}$, $A(7,4) =
A(8,2) = A(11,4) = A(11,8) = A(12,2) = A(12,6) = {0.333333}$,
$A(4,5) = A(4,9) = A(5,3) = A(8,1) = A(8,9) = A(9,3) = A(9,7) =
A(12,1) = A(12,5) = {0.166667}$.\\

The pair of vectors $(x^*, y^*) \in \mathbb{R}^{13 + 13}$ plans given by
\begin{eqnarray*}
  \begin{split}
    x^* &= [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0,
      0.275, 0.725]^T,\\
    y^* &= [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T
    \end{split}
\end{eqnarray*}
is a Nash $(10^{-4},0)$-equlibrium computed in 1500 iterations of
Algorithm  \ref{Tab:algo}. The convergence curves are shown
in Fig \ref{Tab:dgap_curve}. One easy checks that this equilibrium is
feasible. Indeed, one computes \\\\
$E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67
      \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5},
      -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$
and
$E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
  \times 10^{-6}, -1.50 \times 10^{-6},
      2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T$.\\\\
Finally, one checks that ${x^*}^TAy^* = {-0.05555}$,
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of
 the game across iterations are shown in Figure \ref{Tab:dgap_curve}.
We have not benchmarked this against the algorithms proposed in
\cite{nesterov2005a} and Gilpin's et al. \cite{gilpinfirst} because
implementing them for such games would require us to compute the
complicated projections $\prox_{Q_k}$.  We recall that
avoiding these projections was one of the goals of the manuscript.

@@
